---
title: "Project 2 Intermediate Deliverable 1 "
author: "Group 15 Analysts R Us"
date: "2/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Part 1 Data Set Introduction

We found our data set in the Kaggle data base.According to the information from Kaggle, we know that the data set is originally from the National Institute of Diabetes and Digestive and Kidney Diseases.The diagnostic, binary-valued variable investigated is whether the patient shows signs of diabetes according to World Health Organization criteria and all of the patients in this data set lives near Phoenix, Arizona, USA. 

Let us first load the data set. 
```{r}
diabetes <- read.csv("diabetes.csv")
summary(diabetes)
str(diabetes)
```
There are a total of 9 columns in the data set:

`Pregnancies`: Number of times pregnant

`Glucose`: Plasma glucose concentration a 2 hours in an oral glucose tolerance test

`BloodPressure`: Diastolic blood pressure (mmHg)

`SkinThickness`: Triceps skin fold thickness (mm)

`Insulin`: 2-Hour serum insulin (mu U/ml)

`BMI`: Body mass index (weight in kg/(height in m)^2)

`DiabetesPedigreeFunction`: Diabetes pedigree function (a function which scores the likelihood of diabetes based on family history)

`Age`: Age (years)

`Outcome`: Class variable (0 or 1), this is also going to be our response variable

`Business Question` Does a patient have diabetes? 

Throughout this project we will generate the best prediction model based on our data set: predict whether a person has diabetes with information about their number of times pregnant, glucose level, blood pressure, triceps skin fold thickness, insulin, BMI, diabetes pedigree function, and age. If we can predict diabetes based on the above information through a questionnaire, we can target people specifically with high risk of having diabetes if we are a company selling health supplements for diabetes prevention. 

We will first clean the data 
```{r}
#convert number of pregnancy to factors
diabetes$Pregnancies <- as.factor(diabetes$Pregnancies)
#convert outcome to factors 
diabetes$Outcome <- as.factor(diabetes$Outcome)
#divide age into few categories: 21-30 as young adults, 31-45 adults, 46-59 old adults,60+ senior
age_level <- function(x) {
  if(x >= 21 & x <= 30 ) {
    age_level <- "Young adult"
  } else if(x > 30 & x <= 45) {
    age_level <- "Adult"
  } else if(x > 45 & x <= 59) {
    age_level <- "Old adult"
  } else {
    age_level <- "Senior"
  }
  age_level
}
#create a new column in the data set called age level 
#applying the age level function to the Age column
diabetes$age_level <- sapply(diabetes$Age, FUN=age_level) 
diabetes$age_level <- as.factor(diabetes$age_level)
str(diabetes)
summary(diabetes)
```
Now that we finished cleaning up the data, we are going to do some primary exploration on the cleaned data set before we start making and improving prediction models using ANN,kNN, and logistic regression. 

We will first explore the characteristics of people who are diabetic
```{r}
y <- diabetes[diabetes$Outcome == 1, ]

summary(y$Glucose)
summary(y$BloodPressure)
summary(y$SkinThickness)
summary(y$Insulin)
summary(y$BMI)
summary(y$Age)
summary(as.numeric(y$Pregnancies))
summary(as.numeric(y$DiabetesPedigreeFunction))
```
Based on the data, we can see that people with diabetes have glucose levels on average around 141.3, blood pressure around 70.82 mmHg, skin thickness around 22.16 mm, insulin level around 100.3 mu U/ml, BMI around 35.14, and around 37 years of age. The average number of times of pregnancy is around 5. The average score of the likelihood of diabetes based on family history is 0.55. Note that for blood pressure and skin thickness, the median is a lot greater than the mean, indicating that the distribution is negatively skewed, with extremely low values pulling the mean down. 

Next we will explore the characteristics of people who are not diabetic
```{r}

n <- diabetes[diabetes$Outcome == 0, ]

summary(n$Glucose)
summary(n$BloodPressure)
summary(n$SkinThickness)
summary(n$Insulin)
summary(n$BMI)
summary(n$Age)
summary(as.numeric(n$Pregnancies))
summary(as.numeric(n$DiabetesPedigreeFunction))
```
Based on the data, we can see that people without diabetes have glucose levels on average around 110, blood pressure around 68.18 mmHg, skin thickness around 19.66 mm, insulin level around 68.79 mu U/ml, BMI around 30.3, and around 31 years of age. The average number of times of pregnancy is around 4. The average score of the likelihood of diabetes based on family history is 0.43. Note that for blood pressure and skin thickness, the median is greater than the mean, indicating that the distribution is negatively skewed, with extremely low values pulling the mean down. Also for insulin levels, the median is a lot lower than the mean, indicating that there are very extreme high values that pull the mean higher towards to right. The median for age is also lower than the mean, indicating there are people who are older than the majority. 